
---
title: 'Bivariate Visualization'
---

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
```


## Exercise 0: {.unnumbered} 
```{r}
# Load data
elections <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")

# Check it out
head(elections)
```
```{r}
ggplot(elections, aes(x=winner_20))+
  geom_bar()
elections %>%
  count(winner_20)
```
The republicans won about 82.8% of the counties. 

```{r}
ggplot(elections, aes(x=repub_pct_20))+
  geom_histogram()
```
I don't understand the significance of this being negatively skewed. 


## Exercise 1: {.unnumbered} 
```{r}
ggplot(elections, aes(x=repub_pct_16, y=repub_pct_20))+ 
  geom_point()
```
 
## Exercise 2: {.unnumbered} 

```{r}
# This shows the same plot as above in exercise 2, but with crosses. 
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point(color = 'orange', shape=3)+
  geom_text(aes(label = state_abbr))
```

## Exercise 3: {.unnumbered} 

With an increase in republican percentage in 2016, there is an increase in republican percentage in 2020. This relationship is very strong, and it is a positive relationship. A lot of counties in Texas seem to lie above the trend (positive residuals), which could be because they are more republican than the average state. 

## Exercise 4: {.unnumbered} : 
```{r}
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_smooth()
```
```{r}
ggplot(elections, aes(y = repub_pct_20, x = repub_pct_16)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
# Scatterplot of repub_pct_20 vs median_rent
ggplot(elections, aes(x = median_rent, y = repub_pct_20)) +
  geom_point() 


# Scatterplot of repub_pct_20 vs median_age
ggplot(elections, aes(x = median_age, y = repub_pct_20)) +
  geom_point() 

```

Median rent seems to explain repub_pct_20 better: as median rent rises, repub_pct_20 seems to fall, suggesting a moderate negative relationship. Median age does not seem to explain repub_pct_20 very well. 

## Exercise 6: {.unnumbered} 

```{r}
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_point()
```
A scatterplot is very useful for 2 quantitative variables. Historical seems to be a categorical variable with 3 categories, so a boxplot would be a better idea. 

## Exercise 7: {.unnumbered} 

```{r}
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_violin()
```

```{r}
ggplot(elections, aes(y = repub_pct_20, x = historical)) +
  geom_boxplot()
```
This tells us that if a county was historically democrat, people voted less republican in 2020 than if the country was historically purple, which is less than if the county was historically red. 

## Exercise 8: {.unnumbered} 

```{r}
ggplot(elections, aes(x = repub_pct_20, fill=historical)) +
  geom_density(alpha=0.5)
```

## Exercise 9: {.unnumbered} : 

```{r}
# Name two "bad" things about this plot
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density()
```
1. It is opaque, so I can't see part of the data (for the green and red plots). 
2. The colors do not correspond to the colors they are representing (red is blue, blue is red)

```{r}
# What does scale_fill_manual do?
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density() +
  scale_fill_manual(values = c("blue", "purple", "red"))
```
This fixes the second issue with this! The colors correspond to the category they are representing. 

```{r}
# What does alpha = 0.5 do?
# Play around with different values of alpha, between 0 and 1
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = c("blue", "purple", "red"))
```
This fixes the first issue I had with the plot! It increases its transparency. 

```{r}
# What does facet_wrap do?!
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_density() +
  scale_fill_manual(values = c("blue", "purple", "red")) +
  facet_wrap(~ historical)
```
It makes each of the categories into separate small graphs, while using the same axes. 

```{r}
# Let's try a similar grouping strategy with a histogram instead of density plot.
# Why is this terrible?
ggplot(elections, aes(x = repub_pct_20, fill = historical)) +
  geom_histogram(color = "white") +
  scale_fill_manual(values = c("blue", "purple", "red"))
```
This uses a histogram, which is difficult to understand because there is so much information. 

## Exercise 10: {.unnumbered}  

- Which do you like best?
Side-by-side boxplots
- What is one pro of density plots relative to boxplots?
It gives you a visual representation of how skewed the data is
- What is one con of density plots relative to boxplots?
It does not give you any numerical summaries (IQR, median)

## Exercise 11: {.unnumbered} 

```{r}
ggplot(elections, aes(x=historical, fill=winner_20))+
  geom_bar()
```
```{r}
ggplot(elections, aes(x=winner_20, fill=historical))+
  geom_bar()+
  scale_fill_manual(values = c("blue", "purple", "red"))+
  facet_wrap(~ historical)
```

## Exercise 12: {.unnumbered} 

```{r}
# A stacked bar plot
# How are the "historical" and "winner_20" variables mapped to the plot, i.e. what roles do they play?
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar()
```
```{r}
# A faceted bar plot
ggplot(elections, aes(x = winner_20)) +
  geom_bar() +
  facet_wrap(~ historical)
```

```{r}
# A side-by-side bar plot
# Note the new argument to geom_bar
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar(position = "dodge")
```

```{r}
# A proportional bar plot
# Note the new argument to geom_bar
ggplot(elections, aes(x = historical, fill = winner_20)) +
  geom_bar(position = "fill")
```
Part a: An advantage to using a proportional bar chart is that it gives us a proportion of the numbers who voted democrat or republican within each historical category, instead of the raw number. A disadvantage to using a proportional bar chart is that you don't know the raw count of the people who voted democrat or republican within each category- we loose data, because we don't know which category voted most frequently, just the proportion with which they voted. 

Part b: I like side-by-side bar charts (the third graph) because it is easy to follow and understand, the categories are grouped together, and I can see the raw count of votes for each party. 

## Exercise 13: {.unnumbered} 

```{r}
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv")
```

```{r}
# How do 3pm temperatures (temp3pm) differ by location?
ggplot(weather, aes(x=location, y=temp3pm))+ 
  geom_boxplot()
```
The location variable is categorical: there are 3 categories. The temp3pm variable is quantitative. Hence, a boxplot or violin plot would be best at representing this data. 

```{r}
# How might we predict the 3pm temperature (temp3pm) by the 9am temperature (temp9am)?
ggplot(weather, aes(x=temp9am, y=temp3pm))+
  geom_point()+
  geom_smooth()
```
Both of these variables are quantitative, so a scatterplot would be best 

```{r}
# How do the number of rainy days (raintoday) differ by location?
ggplot(weather, aes(x=location, fill=raintoday))+ 
  geom_bar()
```
Both are categorical variables. 
