---
title: "Chapter 5"
---
```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
```


```{r}
weather <- read.csv("https://mac-stat.github.io/data/weather_3_locations.csv") %>% 
  mutate(date = as.Date(date)) 
```

To make a stacked bar chart:

```{r}
woll <- weather %>%
  filter(location == "Wollongong") %>% 
  mutate(date = as.Date(date))  
```


```{r}
# Default: stacked bars
ggplot(woll, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar()
```

To make side-by-side bar charts: 

```{r}
# Side-by-side bars
ggplot(woll, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar(position = "dodge")
```

Proportional bar charts: 

```{r}
# Proportional bars
# position = "fill" refers to filling the frame, nothing to do with the color-related fill
ggplot(woll, aes(x = raintoday, fill = raintomorrow)) + 
  geom_bar(position = "fill")
```

To visualize 3 variables: 

- Play with alpha 
- Play with colors 
- Play with facets 
- Add an axis 

If you have 3 variables, how do you represent: 
- 3 categorical variables: Use a stacked bar chart with color, with facets 
- 2 categorical and 1 numerical: Use a stacked density plot with color, with facets (or histogram, density, violin)
- 2 numerical and 1 categorical: Use a scatterplot with color or facets (categorical) (or heatmap or star)
- 3 numerical: Use a scatterplot with a discrete/ binned version of the 3rd numerical variable and change the color, shape, size, or transparency 
